 

### 2021-07-27 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd

**Similar to photosynthesis, respiration tends to be higher in canopy sun-leaves [Table 2, @chenLeafEconomicsSpectrum2020], but its temperature sensitivity ($Q_{10}$) shows no definite trend along the vertical gradient [@bolstad_foliar_1999; Weerasinghe et al.2014].** Specifically, the $Q_{10}$ of respiration can be constant within vertical profiles and in seedling sun vs. shade leaves [@@carterExperimentalWarmingTropical; @weerasingheCanopyPositionAffects2014; @bolstad_foliar_1999; @zaragoza-castellsDoesGrowthIrradiance2007; @zaragoza-castellsClimatedependentVariationsLeaf2008; @xuSeasonalVariationTemperature2006], greater in upper-canopy leaves [@turnbullScalingFoliarRespiration2003; @harleyEffectsLightTemperature1996], or greater in the lower canopy [@griffinCanopyPositionAffects2002]. Recent observations suggest that tropical understory leaves down-regulate $Q_{10}$ with rising temperature [@carterExperimentalWarmingTropical]; this may vary based on leaf age and leaf traits, as in evergreen species, $Q_{10}$ has shown to be greater in younger than older leaves [@zhouLeafageEffectsTemperature2015].

Further, we would expect $Q_{10}$ to also vary with soil moisture based on the evidence of canopy-position effect gradients in respiration rates for *Quercus rubra* being much greater in mesic than in xeric sites [@xuSeasonalVariationTemperature2006].

@@

### Tree metabolism, growth, and survival

>>>>

**Similar to photosynthesis, respiration tends to be higher in canopy sun-leaves [Table 2, @chenLeafEconomicsSpectrum2020], but its temperature sensitivity ($Q_{10}$) shows no definite trend along the vertical gradient [@bolstad_foliar_1999; Weerasinghe et al.2014].** Specifically, the $Q_{10}$ of respiration can be constant within vertical profiles and in seedling sun vs. shade leaves [@@carterExperimentalWarmingTropical; @weerasingheCanopyPositionAffects2014; @bolstad_foliar_1999; @zaragoza-castellsDoesGrowthIrradiance2007; @zaragoza-castellsClimatedependentVariationsLeaf2008; @xuSeasonalVariationTemperature2006], greater in upper-canopy leaves [@turnbullScalingFoliarRespiration2003; @harleyEffectsLightTemperature1996], or greater in the lower canopy [@griffinCanopyPositionAffects2002].This may vary based on forest type [**REF**], leaf traits and leaf age, as in evergreen species, $Q_{10}$ has shown to be greater in younger than older leaves [@zhouLeafageEffectsTemperature2015]. Recent observations suggest that tropical understory leaves down-regulate $Q_{10}$ with rising temperature [@carterExperimentalWarmingTropical]. Further, we would expect $Q_{10}$ to also vary with soil moisture based on the evidence of canopy-position effect gradients in respiration rates for *Quercus rubra* being much greater in mesic than in xeric sites [@xuSeasonalVariationTemperature2006].

@@

### Tree metabolism and growth, and survival
